Variants on the Flight and Maintenance Planning problem

### 2018-07-24 latex/notes/model/model.tex
Finished adding remaining constraints to model.

    \begin{tabular}{p{5mm}p{170mm}}
        $\mathcal{I}$     &  resources. \\
        $\mathcal{T}$     &  time periods inside planning horizon. \\
        $\mathcal{J}$     &  tasks. \\
        $\mathcal{K}$     &  cluster of resources that share the same capacities to do tasks. \\
        $\mathcal{T}_j$     &  time periods $t \in \mathcal{T}$ in which task $j$ is active. \\
        $\mathcal{J}_t $    &  tasks $j \in \mathcal{J}$ to be realized in period $t$. \\
        $\mathcal{I}_j$     &  resources $i \in \mathcal{I}$ that can be assigned to task $j$. \\
        $\mathcal{I}_k$     &  resources $i \in \mathcal{I}$ that are included in cluster $k$. \\
        $\mathcal{O}_i$     &  tasks $j \in \mathcal{J}$ for which resource $i$ can be used. \\
        $\mathcal{T}^{s}_t$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - M+1\}},  ..., {t}\}$. \\
        % $\mathcal{T}^{st}_t$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - S+1\}},  ..., {t}$\}. \\
        $\mathcal{T}^{m}_t$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - E^{m}+1\}},  ..., {t}\}$. \\
        $\mathcal{T}^{M}_t$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - E^{M}+1\}},  ..., \max{\{1, t - E^{m}+1\}}$\}. \\
        $\mathcal{T}^{MT}_{jt}$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - MT_j\}},  ..., {t}\}$. \\
        % $\mathcal{T}^{M}_t$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - E^{M}+1\}},  ..., \max{\{1, t - E^{m}+1\}}$\}. \\
    \end{tabular}

@@

  % TODO: previously assigned missions.
  % TODO: previous ret.
  % TODO: previous maintenaces.

    \begin{tabular}{p{8mm}p{167mm}}
        $H_j$             & amount of resource time required by task $j$. \\
        $R_j$             & number of resources required by task $j$. \\
        $MT_j$            & minimum number of consecutive periods a resource can be assigned to task $j$. \\
        % $aut_i$           & maximal absolute usage time for resource $i$. \\
        % $aet_i$           & maximal absolute elapsed time for resource $i$. \\
        $M$               & maintenance duration in number of periods. \\
        $M^{max}$         & maximum number of simultaneous maintenance operations. \\
        % $E$               & remaining elapsed time after a maintenance. \\
        $E^{M}$         & maximum number of periods between two consecutive maintenances. \\
        $E^{m}$         & minimum number of periods between two consecutive maintenances. \\
        $H$               & remaining usage time after a maintenance. \\
        % $S$               & Storage duration in number of periods. \\
        % $W_1$             & weight of the first objective in the objective function. \\
        % $W_2$             & weight of the second objective in the objective function. \\
        % $N_t$             & number of resources in already-planned maintenances in period $t$ at the beginning of the planning horizon.\\
        % $D_t$             & number of resources to be assigned in total in period $t$. \\
        $AK_{kt}$         & maximum number of resources in cluster $k$ that can be simultaneously under maintenance at period $t$.\\
        $Rut^{Init}_{i}$  & remaining usage time for resource $i$ at the start of the planning horizon. \\
        $Ret^{Init}_{i}$  & remaining elapsed time for resource $i$ at the start of the planning horizon. \\
        % $A_j$     &  resources $i \in \mathcal{I}$ that can be assigned to task $j$. \\
        % $Ret^{Init}_{sum}$& sum of remaining elapsed times at the start of the planning horizon. \\
        % $Rut^{Init}_{sum}$& sum of remaining elapsed time at the start of the planning horizon. \\
    \end{tabular}

@@

    where the first expression counts all the resource hours given via maintenances and the second expression measures the amount of remaining hours at all resources at the end of the planning horizon.

    where we want to minimize the total number of resource hours given via maintenances and we want to maximize the final state measured in available hours of resources. These two objectives have the same units and can be easily compared and they ensure the resource hours are used in the most efficient way possible.
    % where weights $W_1$ and $W_2$ are chosen by the decision maker. $W_1$ penalizes the maximum number of resources in maintenance among all periods. $W_2$ penalizes the maximum number of unavailable resources among all periods.
    The following constraints are used in the model:       
    \begin{align}
        % maximum capacity1:
        % & \sum_{t' \in \mathcal{T}^{s}_t} \sum_{i \in \mathcal{I}} m_{it'} + N_t \leq m_{max}
        %   & t \in \mathcal{T} \label{eq:capacity1}\\
        & \sum_{t \in \mathcal{T}, i \in \mathcal{I}} m_{it} \leq M^{max} \label{eq:capacity1}\\
        % availability
       % & \sum_{t' \in \mathcal{T}^{s}_t} \sum_{i \in \mathcal{I}} m_{it'} + N_t + D_t\leq u_{max} 
       %  &t \in \mathcal{T} \label{eq:avalaibility1}\\
        % min assignments:
        & \sum_{i \in \mathcal{I}_j} a_{jti} = R_j
                & j \in \mathcal{J}, t \in \mathcal{T}_j  \label{eq:taskres}\\
        % just doing one thing at any given time:
        & \sum_{t' \in \mathcal{T}^{s}_t} m_{it'} + \sum_{j \in \mathcal{J}_t \cap \mathcal{O}_i} a_{jti} \leq 1 
                & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:state}
    \end{align}

@@

    \begin{align}
        % if start task => assign start
        & a^s_{jti} \geq a_{jti} - a_{j(t-1)i} 
                & j \in \mathcal{J}, t \in \mathcal{T}_j, i \in \mathcal{I}_j \label{eq:start1} \\
      % # if we start a task, we need at least X periods of tasks assignments
        & a^s_{jt'i} \leq a_{jti}
        & j \in \mathcal{J}, t \in \mathcal{T}_j, i \in \mathcal{I}_j, t' \in \mathcal{T}^{MT}_{jt} \label{eq:start2}
    \end{align}

    Constraints (\ref{eq:start1}) controls the definition of starting a consecutive assignment: the variable activates if a resource $i$ is assigned a task $j$ in period $t$ and was not assigned this task in the previous period ($t-1$).
    Constraints (\ref{eq:start2}) controls the minimum duration of a consecutive task assignment. If an assignment to task $j$ is started in period $t$, it needs to continue in the following $t' \in \mathcal{T}^{MT}_{jt}$ periods.

@@

    Constraints (\ref{eq:avalaibility-cluster}) control a minimum availability for each cluster $k$ of resources. A cluster is defined by the most general group of resources that is required, exclusively, by at least one task.

    \begin{align}
        % remaining used time
         & rut_{it} \leq rut_{i(t-1)} + H m_{it} - \sum_{j \in \mathcal{J}_t \cap \mathcal{O}_i} a_{jti} H_j & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:rut_upper}\\
        & rut_{i0} = Rut^{Init}_i
               & i \in \mathcal{I} \label{eq:rut_initial}\\
        & rut_{it} \geq H m_{it}
                & t \in \mathcal{T}, i \in \mathcal{I}\label{eq:rut_lower}\\ 
        & rut_{it} \in [0,H]
                & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:mu} \\               
        % & ret_{it} \leq ret_{it-1} - 1 + \sum_{t' \in \mathcal{T}^{st}_t} s_{it'} + E m_{it}
        %         & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:ret_upper}\\
        % & ret_{i0} = Ret^{Init}_i
        %         & i \in \mathcal{I} \label{eq:ret_initial}\\
        % & ret_{it} \geq E m_{it}
        %         & t \in \mathcal{T}, i \in \mathcal{I}\label{eq:ret_lower}\\                 
        % & ret_{it} \in [0,E]
        %         & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:me}\\
        % & \sum_{i \in \mathcal{I}} ret_{it} \geq Ret^{Init}_{sum}
        %       & t = |\mathcal{T}| \label{eq:min_ret}\\
        % & \sum_{i \in \mathcal{I}} rut_{i|\mathcal{T}|} \geq Rut^{Init}_{sum}\\
        & m_{it'} + m_{it} \leq 1
          & t, \in \mathcal{T}, t' \in \mathcal{T}^{m}_t, i \in \mathcal{I}\label{eq:ret_min}\\ 
        & \sum_{t' \in \mathcal{T}^{M}_t} m_{it'} \leq  m_{it}
          & t, \in \mathcal{T}, i \in \mathcal{I}\label{eq:ret_max}
    \end{align}

    % \begin{align}
    % & m_{it'} + m_{it} \leq 1
    %   & t, \in \mathcal{T}, t' \in \mathcal{T}^{m}_t, i \in \mathcal{I}\label{eq:ret_better}\\ 
    % & \sum_{t' \in \mathcal{T}^{M}_t - \mathcal{T}^{m}_t} m_{it'} \leq  m_{it}
    %   & t, \in \mathcal{T}, i \in \mathcal{I}\label{eq:ret_better}\\ 
    % \end{align}

@@

    The minimum and maximum elapsed time are defined by (\ref{eq:ret_min}) and (\ref{eq:ret_max}) respectively. Constraints (\ref{eq:ret_min}) guarantee that if a maintenance is done in some period $t$, we know that it cannot be done in the following $t' \in \mathcal{T}^{m}_t$ periods. Constraints (\ref{eq:ret_max}) guarantee that, if we have a maintenance in period $t$, we have to start at least one maintenance among periods $t' \in \mathcal{T}^{M}_t}$.

>>>>

    \begin{tabular}{p{5mm}p{170mm}}
        $\mathcal{I}$     &  resources. \\
        $\mathcal{T}$     &  time periods inside planning horizon. \\
        $\mathcal{J}$     &  tasks. \\
        $\mathcal{K}$     &  cluster of resources that share the same capacities to do tasks. \\
        $\mathcal{T}_j$     &  time periods $t \in \mathcal{T}$ in which task $j$ is active. \\
        $\mathcal{J}_t $    &  tasks $j \in \mathcal{J}$ to be realized in period $t$. \\
        $\mathcal{I}_j$     &  resources $i \in \mathcal{I}$ that can be assigned to task $j$. \\
        $\mathcal{I}_k$     &  resources $i \in \mathcal{I}$ that are included in cluster $k$. \\
        $\mathcal{O}_i$     &  tasks $j \in \mathcal{J}$ for which resource $i$ can be used. \\
        $\mathcal{T}^{s}_t$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - M+1\}},  ..., {t}\}$. \\
        % $\mathcal{T}^{st}_t$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - S+1\}},  ..., {t}$\}. \\
        $\mathcal{T}^{m}_t$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ {t}, ..., \min{\{|\mathcal{T}|, t + E^{m}-1\}}\}$. \\
        $\mathcal{T}^{M}_t$ &  time periods $t' \leq |\mathcal{T}| - E^{M}$ such that $t' \in \{ t + E^{m}-1 , ...,  t + E^{M}-1 \}$. \\
        $\mathcal{T}^{m_{ini}}_i$ &  time periods $t \in \mathcal{T}$ such that $t \in \{ 1, ..., \max{\{0, Ret^{Init}_{i} - E^M + E^m \}}\}$. \\
        $\mathcal{T}^{M_{ini}}_i$ &  time periods $t \in \mathcal{T}$ such that $t \in \{ \max{\{0, Ret^{Init}_{i} - E^M + E^m \}} , ...,  Ret^{Init}_{i} \}$. \\
        $\mathcal{T}^{MT}_{jt}$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - MT_j\}},  ..., {t}\}$. \\
    \end{tabular}

@@

  % TODO: previously assigned missions.

    \begin{tabular}{p{8mm}p{167mm}}
        $H_j$             & amount of resource time required by task $j$. \\
        $R_j$             & number of resources required by task $j$. \\
        $MT_j$            & minimum number of consecutive periods a resource can be assigned to task $j$. \\
        % $aut_i$           & maximal absolute usage time for resource $i$. \\
        % $aet_i$           & maximal absolute elapsed time for resource $i$. \\
        $M$               & maintenance duration in number of periods. \\
        $M^{max}$         & maximum number of simultaneous maintenance operations. \\
        % $E$               & remaining elapsed time after a maintenance. \\
        $E^{M}$         & maximum number of periods between two consecutive maintenances. \\
        $E^{m}$         & minimum number of periods between two consecutive maintenances. \\
        $H$               & remaining usage time after a maintenance. \\
        % $S$               & Storage duration in number of periods. \\
        % $W_1$             & weight of the first objective in the objective function. \\
        % $W_2$             & weight of the second objective in the objective function. \\
        % $N_t$             & number of resources in already-planned maintenances in period $t$ at the beginning of the planning horizon.\\
        % $D_t$             & number of resources to be assigned in total in period $t$. \\
        $AK_{kt}$         & maximum number of resources in cluster $k$ that can be simultaneously under maintenance at period $t$.\\
        $Rut^{Init}_{i}$  & remaining usage time for resource $i$ at the start of the planning horizon. \\
        $Ret^{Init}_{i}$  & remaining elapsed time for resource $i$ at the start of the planning horizon. \\
        $A^{init}_{jti}$  & =1 if resource $i \in \mathcal{I}$ has task $j$ pre-assigned in period $t$. \\
        % $Ret^{Init}_{sum}$& sum of remaining elapsed times at the start of the planning horizon. \\
        % $Rut^{Init}_{sum}$& sum of remaining elapsed time at the start of the planning horizon. \\
    \end{tabular}

@@

    where the first expression counts all the usage hours given to resources via maintenances and the second expression measures the amount of remaining usage hours in the resources at the end of the planning horizon.

    We want to minimize the total number of maintenances and we want to maximize the final state measured. These two objectives have the same units, can be easily compared and they ensure the resource hours are used in the most efficient way possible.

    The following constraints are used in the model:       
    \begin{align}
        % maximum capacity1:
        % & \sum_{t' \in \mathcal{T}^{s}_t} \sum_{i \in \mathcal{I}} m_{it'} + N_t \leq m_{max}
        %   & t \in \mathcal{T} \label{eq:capacity1}\\
        & \sum_{t \in \mathcal{T}, i \in \mathcal{I}} m_{it} \leq M^{max} \label{eq:capacity1}\\
        % availability
       % & \sum_{t' \in \mathcal{T}^{s}_t} \sum_{i \in \mathcal{I}} m_{it'} + N_t + D_t\leq u_{max} 
       %  &t \in \mathcal{T} \label{eq:avalaibility1}\\
        % min assignments:
        & \sum_{i \in \mathcal{I}_j} a_{jti} = R_j
                & j \in \mathcal{J}, t \in \mathcal{T}_j  \label{eq:taskres}\\
        % just doing one thing at any given time:
        & \sum_{t' \in \mathcal{T}^{s}_t} m_{it'} + \sum_{j \in \mathcal{J}_t \cap \mathcal{O}_i} a_{jti} \leq 1 
                & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:state}
    \end{align}

@@

    \begin{align}
        % if start task => assign start
        & a^s_{jti} \geq a_{jti} - a_{j(t-1)i} - A^{init}_{j(t-1)i}
                & j \in \mathcal{J}, t \in \mathcal{T}_j, i \in \mathcal{I}_j \label{eq:start1} \\
      % # if we start a task, we need at least X periods of tasks assignments
        & a^s_{jt'i} \leq a_{jti}
        & j \in \mathcal{J}, t \in \mathcal{T}_j, i \in \mathcal{I}_j, t' \in \mathcal{T}^{MT}_{jt} \label{eq:start2}
    \end{align}

    Constraints (\ref{eq:start1}) control the definition of starting a consecutive assignment: the variable is active if a resource $i$ is assigned a task $j$ in period $t$ and was not assigned this same task in the previous period ($t-1$). The previous period can include the period before the start of the planning horizon ($A^{init}_{j(t-1)i}$ parameter).
    Constraints (\ref{eq:start2}) control the minimum duration of a consecutive task assignment. If an assignment to task $j$ is started in period $t$, it needs to continue for the following $t' \in \mathcal{T}^{MT}_{jt}$ periods.

@@

    Constraints (\ref{eq:avalaibility-cluster}) guarantee a minimum availability for each cluster $k$ of resources. A cluster is defined by the biggest group of resources that is required, exclusively, by at least one task.

    \begin{align}
        % remaining used time
         & rut_{it} \leq rut_{i(t-1)} + H m_{it} - \sum_{j \in \mathcal{J}_t \cap \mathcal{O}_i} a_{jti} H_j & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:rut_upper}\\
        & rut_{i0} = Rut^{Init}_i
               & i \in \mathcal{I} \label{eq:rut_initial}\\
        & rut_{it} \geq H m_{it}
                & t \in \mathcal{T}, i \in \mathcal{I}\label{eq:rut_lower}\\ 
        & rut_{it} \in [0,H]
                & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:mu} \\               
        % & ret_{it} \leq ret_{it-1} - 1 + \sum_{t' \in \mathcal{T}^{st}_t} s_{it'} + E m_{it}
        %         & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:ret_upper}\\
        % & ret_{i0} = Ret^{Init}_i
        %         & i \in \mathcal{I} \label{eq:ret_initial}\\
        % & ret_{it} \geq E m_{it}
        %         & t \in \mathcal{T}, i \in \mathcal{I}\label{eq:ret_lower}\\                 
        % & ret_{it} \in [0,E]
        %         & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:me}\\
        % & \sum_{i \in \mathcal{I}} ret_{it} \geq Ret^{Init}_{sum}
        %       & t = |\mathcal{T}| \label{eq:min_ret}\\
        % & \sum_{i \in \mathcal{I}} rut_{i|\mathcal{T}|} \geq Rut^{Init}_{sum}\\
    \end{align}

@@

    \begin{align}
        & m_{it'} + m_{it} \leq 1
          & t \in \mathcal{T}, t' \in \mathcal{T}^{m}_t, i \in \mathcal{I}\label{eq:ret_min}\\ 
        & \sum_{t' \in \mathcal{T}^{M}_t} m_{it'} \geq  m_{it}
          & t \in \mathcal{T}, i \in \mathcal{I}\label{eq:ret_max}\\
        & m_{it} = 0
          & t \in \mathcal{T}^{m_{ini}}_i, i \in \mathcal{I}\label{eq:ret_min_init} \\
        & \sum_{t \in \mathcal{T}^{M_{ini}}_i} m_{it} \geq  1 
          & i \in \mathcal{I}\label{eq:ret_max_init}
    \end{align}

    The minimum and maximum elapsed time are defined by (\ref{eq:ret_min}) and (\ref{eq:ret_max}) respectively. Constraints (\ref{eq:ret_min}) guarantee that if a maintenance is done in some period $t$, we know that it cannot be done in the immediately consecutive $t' \in \mathcal{T}^{m}_t$ periods. Constraints (\ref{eq:ret_max}) guarantee that, if we have a maintenance in period $t$, we need to start at least one maintenance among periods $t' \in \mathcal{T}^{M}_t$. Constraints (\ref{eq:ret_min_init}) and (\ref{eq:ret_max_init}) control the minimum and maximum remaining elapsed time at the beginning of the planning period, respectively. They follow the same rules as the previous constraints for the particular case of the start of the planning horizon.
